---
Author: Toto Kamari
Date: 9/15/2020
---

Loading and Preprocessing the Data
```{r}
a <- read.csv("./activity.csv")
```

===============================================================

What is the mean total number of steps taken per day?
Making a histogram of the total number of steps taken each day:
```{r}
b <- aggregate(a["steps"], by=list(a$date), sum, na.rm=TRUE)
hist(b$steps, main = "Total Steps Taken per Day", xlab = "Steps")
```

Calculating the mean and median of the total number of steps taken per day:
```{r}
c <- mean(b$steps)
d <- median(b$steps)
```
The mean number of steps per day is 'r c'
The median number of steps per day is 'r d'

================================================================

What is the average daily activity pattern?
Making a time series plot of the average number of steps per 5 minute interval
```{r}
e <- aggregate(a["steps"], by=list(a$interval), mean, na.rm=TRUE)
plot(e$Group.1, e$steps, type = "l", main = "Average Steps per 5 minute Intervals",
     xlab = "5-minute Interval", ylab = "Average Number of Steps")
```

Finding what 5-minute interval has the maximum number of steps
```{r}
f <- e$Group.1[which.max(e$steps)]
```
The 5-minute interval that contains the maximum number of steps is 'r f'

=================================================================

Imputing missing values
Calculating the total number of missing values
```{r}
g <- is.na(rowSums(a[,c(1,3)]))
h <- sum(g)
```
The number of NAs is 'r h'

Replacing the NAs with the interval's mean
```{r}
colnames(e) <- c("interval", "steps")
i <- merge(x = a, y = e, by = "interval", suffixes = c("", "Average"))
j <- i[with(i, order(date)), ]
j$steps[g] <- j$stepsAverage[g]
k <- j[,1:3]
```

Making a new histogram
```{r}
l <- aggregate(k["steps"], by=list(k$date), sum)
hist(l$steps, main = "Total Steps Taken per Day", xlab = "Steps")
```

Calculating the new mean and median
```{r}
m <- mean(l$steps)
n <- median(l$steps)
```
The new mean number of steps taken per day is 'r m'
The new median number of steps taken per day is 'r n'

==================================================================

Are there differences in activity patterns between weekdays and weekends?
Making a panel plot to show averages across weekdays and weekends
```{r}
library(plyr)
library(lattice)
k$date <- as.POSIXlt(k$date)
o <- cbind(k, (weekdays(k$date) == "Sunday" | weekdays(k$date) == "Saturday"))
colnames(o) <- c("interval", "steps", "date", "daytype") 
p <- ddply(o, c("interval", "daytype"), summarize, outVal = mean(steps))
p$daytype <- as.factor(p$daytype)
levels(p$daytype <- c("Weekends", "Weekdays"))
xyplot(outVal ~ interval | daytype, data = p, type = "l", layout = c(1,2),
       ylab = "Number of Steps", xlab = "Interval")
```
